Freezing of a Weeks-Chandler-Anderson Fluid using Classical Density Functional Theory

### 2021-01-12 papers/thesis-kirstie/thesis.tex
edits

Classically, fluids have been modeled by a collection of small, 
impenetrable spheres colliding with one another like billiard balls. 
In the simplest case, the spheres are treated as having only a repulsive 
potential. This model, called the hard-sphere fluid, has served as the 
primary reference system for fluids. This is due to the fact that the 
repulsion between atoms is much stronger than their attraction, and so 
the repulsive potential, which is descriptive of hard spheres, dominates 
the interactions between atoms.

In more recent times, other model fluids have been developed.
Zwanzig introduced the square-well 
fluid in 1954 which is a perturbation of the hard-sphere fluid created by adding 
a hard-edge attractive potential to the repulsive potential of the 
hard-sphere fluid~\cite{ZwanzigSqrWell}. In 1964 Rowlinson studied 
spheres with softly fading repulsive potentials which he modeled by making
the radius of the spheres temperature dependent~\cite{rowlinson1964statistical}. 
This reflects the idea that soft-spheres penetrate more deeply into 
each other when they collide at higher temperatures. 
In 1967 Barker and Henderson combined the perturbation approach of Zwanzig 
with Rowlinson's idea of using temperature-dependent radii to model spheres 
with soft repulsive potentials using hard-spheres~\cite{barker1967perturbation}. 
In this model, 
%a soft-sphere fluid is modeled by using a hard-sphere fluid with 
hard spheres are given larger radii 
at low temperatures, and smaller radii at high temperatures. 
In 1971 Weeks, 
Chandler, and Anderson used temperature and density dependent hard-sphere 
radii to model spheres with the more realistic softly fading repulsive 
portion of the Leonard-Jones potential~\cite{andersen1971relationship}, 
which is known as the Weeks-Chandler-Anderson model fluid. 
Models continue to be developed to characterize fluids with particular 
properties, or to address particular problems like surface tension, 
interfaces, adsorption, wetting or confinment.
\textcolor{red}{A review of some model fluids and problems that have been
studied more recently is given by L\"{o}wen~\cite{lowen}.}

While models for particular fluids were being created, powerful methods 
to predict the properties of fluids were also being developed. 
In the late 1950s, Scaled Particle Theory (SPT)~\cite{ReissSPT} was 
introduced from which an equation of state, and an expression for the 
excess Helmholtz free energy of a homogeneous hard-sphere fluid can be 
obtained~\cite{Hansen}. 
In 1958, Percus-Yevick introduced a method to describe the thermodynamics 
of the hard-sphere fluid based on an approximate closure to the 
Ornstein-Zernike integral equation~\cite{PercusYevick}.
%Percus and Yevick took another approach to the hard-sphere fluid.
In 1961 Morita and Hiroike introduced a density functional 
approach~\cite{MoritaDFT} to finding the equilibrium Grand and Helmholtz free energies  
%, and the corresponding equilibrium number density profile 
of a classical inhomogeneous fluid. 
Density functional theory was further developed for inhomogeneous classical fluids
by Stillinger and Buff (1962)~\cite{stillingerandbuff}, Lebowitz and 
Percus (1963)~\cite{lobowitzandpercus}, and Mermin (1965)~\cite{merium}.
To use classical Density Functional Theory (cDFT), an expression for the free energy 
as a functional of a spatially varying average single-particle number density 
%that varies spatially 
is constructed for a particular fluid.
%DFT Evans 1979
%Led by the consistency between results from SPT and Percus-Yevick, 
\textcolor{red}{In 1989 Rosenfeld introduced Fundamental Measure Theory (FMT)~\cite{rosenfeld1989,rosenfeld1988} 
to form a functional for the hard-sphere fluid based on the idea that 
%the lowest term in the virial expansion of the excess chemical potential 
the lowest order term in the virial expansion of the excess chemical 
potential of a hard-sphere fluid can 
be expressed in terms of a weighted density using a weight function 
%using a weight 
%function 
of the form of 
a backward Heaviside step function}
%\begin{equation}
   %F_{ex}=k_BT\int_{volume} n(\vec r) \theta(|\vec r - \vec r'|-(\vec R_i + \vec R_j)) d\vec r
%\end{equation}
that can itself be expressed in terms of a sum of convolutions involving 
a set of weight functions related to the geometry of a sphere. 
For two spheres with radii, 
$R_i$  and $R_j$, the convolution is written as~\cite{rosenfeld1989}
\begin{displaymath}
   \theta(|\vec r - \vec r'|-(R_i + R_j)) = w_i^{(3)} \otimes w_j^{(0)} + w_j^{(3)} \otimes w_i^{(0)} + w_i^{(1)} \otimes w_j^{(2)} + w_j^{(1)} \otimes w_i^{(2)}\end{displaymath}
\begin{equation}   + \vec w_i^{(1)} \otimes \vec w_j^{(2)} + \vec w_j^{(1)} \otimes \vec w_i^{(2)}
\end{equation}
FMT gives the same result for the excess Helmholtz free energy of the hard-sphere fluid as 
SPT does, from which it can be constructed~\cite{rosenfeld1989,Hansen,santos2012phi3},
and for the direct correlation function approximation given by Percus-Yevick. 
However, it did not predict freezing due to one term in the expression that
diverged when the fluid was highly confined.  
Tarazona later solved this problem by introducing a tensor version of 
one of the energy density terms 
in the functional~\cite{tarazonaphi3, santos2012phi3} 
which took into account the non-spherical, and thus 
non-symmetric, shape of the cavity in which the fluid is confined 
while also yielding the correct excess free energy for the fluid as its 
confinment is reduced from a three-dimensional volume (as in a bulk fluid), 
to a two-dimensional plane (as in a thin-film), to a one-dimensional 
line (as in a fluid confined to a capillary tube), to a zero-dimensional 
dot, as in an atom located at a lattice site within a crystal~\cite{rothreview}.

In 1999 Schmidt adopted FMT for spheres with soft 
interatomic potential energies
(which rise to infinity over a distance, rather than instantaneously 
like the potential energy between hard spheres)
forming Soft Fundamental Measure Theory (SFMT)~\cite{schmidt1999density}. 
%by replacing $w^{(3)}$ with a weight function that would result in  
%soft interatomic potentials (which rise to infinity over a distance, 
%rather than instantaneously).
Schmidt applied his Soft Fundamental Measure Theory to various fluids 
with soft, short-range repulsive and attractive potentials, and developed 
an error function model 
%based on a Gaussian weight function 
for spheres with steep, short-range soft repulsive 
potentials~\cite{schmidt2000fluid}.
%, similar to that exhibited by a WCA fluid. 
\textcolor{red}{For some more recent developments, see the review of new developments in 
classical Density Functional Theory by Evans ~\cite{evansreview} (2016), 
Theory of Simple Fluids by Hansen and McDonald~\cite{Hansen} (2013), 
and a recent generalization of the Barker-Henderson model to inhomogeneous 
fluids by Tschopp, Vuijk, Sharma and Brader~\cite{BHextended} (2020).}

\textcolor{red}{Classical density functional theory has also been applied 
to phase transitions, and freezing, but not until recently has there
been good progress. Ramakrishnan and Yussouff studied freezing by viewing a 
crystal as a highly inhomogeneous liquid~\cite{} (1979). Oettel and fellow 
researchers studied hard-sphere crystals with ~\cite{} (2010). Mortazavifar and Oettel
studied the fluid and crystal phases of the Asakura-Oosawa model fluid (2016). 
Lutsko and Lam studied studied crystallization of the Leonard-Jones fluid~\cite{lutsko} (2018),
and Lutsko and Schoonen looked at cDFT applied to the solid state~\cite{2020}.
In all these studies the hard-sphere fluid has been used as the reference system,
which is implemented with FMT, and to which other terms are added to create 
the full functional for the fluid being studied.}
In this paper, classical Density Functional Theory will be used to predict 
the freezing of a Weeks-Chandler-Anderson model fluid directly by applying 
SFMT with Schmidt's proposed error function model, 
in which we have incorporated our temperature dependent parameters to 
fit the functional to the soft-sphere WCA fluid. 
%form the functional for the excess Helmholtz free energy. 
%for generating the weight functions, and Tarazona's tensor functional. 
Results will be compared to those generated from Monte-Carlo simulations.

%Results will be compared to those from Monte-Carlo simulations.

@@

The repulsive portion of the Lennard-Jones potential forms the potential for a 
Weeks-Chandler-Anderson (WCA) model fluid~\cite{andersen1971relationship},
shown in Figure \ref{fig:WCA_potential}. 
The WCA potential is described by the equation 
\begin{align} \label{eq:VWCA}
    V_{WCA}=\left\{\begin{array}{rcl} {4\epsilon{\left[\left(\frac{\sigma}{r}\right)^{12} - \left(\frac{\sigma}{r}\right)^6 \right]}+\epsilon} & \mbox{for} & 0<r<{2R} \\ 0 & \mbox{for} & r>2R \end{array}\right.
\end{align} 
The WCA potential decreases from infinity as the distance between the 
centers of the atoms r increases, and becomes, and remains, essentially 
zero when r=$2^\frac{1}{6}\sigma=2R\approx{1.122}\sigma$ where R is the 
radius of the sphere. 
As with the Lenard-Jones potential, the repulsive WCA potential increases 
rapidly and approaches infinity as the two spheres begin to overlap. This 
is a more realistic repulsive potential than that of the hard-sphere model 
which goes instantly to infinity when the two spheres make contact (at r=2R), 
and is zero otherwise. Because the WCA spheres have no hard boundary in 
their interatomic potential, they are called ``soft spheres'' and their 
gradually fading potential makes them slightly ``squishy'', as illustrated in
Figure~\ref{fig:TwoSpheres}.

@@

A phase transition occurs along a line on a 
Pressure-Temperature (P-T) phase diagram, such as can be seen in the phase 
diagram shown in Figure~\ref{fig:P-T_Diagram}. 
%
%Points along the phase boundary that separates the solid and the fluid phases 
%indicate a phase transition where the solid and fluid phases coexist, 
%and show that both the solid and the fluid have the same temperature (necessary 
%for thermal equilibrium)
%and the same pressure (necessary for mechanical 
%equilibrium)at a phase transition, and the same chemical potential 
%(necessary for diffusive equilibrium). 
%
This phase diagram is 
representative of some substance 
which exhibits both attraction and repulsion. However, since 
the atoms in a WCA fluid do not exhibit attraction there will not be a 
dividing line between the gas and liquid phases, there will just simply
be a fluid.\textcolor{red}{liquid in this paper.}
The phase diagram indicates the equilibrium state a substance will be in at
different temperatures and pressures.
At a phase tranition, both the fluid and the solid have 
the same temperature (necessary 
for thermal equilibrium), the same pressure (necessary for mechanical 
equilibrium) and the same chemical potential (necessary for diffusive equilibrium).
%The temperature , pressure, and chemical potential will be the same for both
%the fluid and the solid at a fluid-solid phase transition.

The pressure can be related to the Helmholtz free energy F, which is minimized 
for a system at fixed temperature, volume and number of particles, as the 
entropy of the system and its surroundings is maximized. The negative of 
the partial derivative of the Helmholtz free energy with respect to volume V, 
and at a fixed temperature T and number of particles N, equals the pressure. 
\begin{equation}{P=-\frac{\partial{F}}{\partial{V}}\bigg|_{T,N}}\end{equation}
\noindent Dividing both the Helmholtz free energy and the volume by the 
number of particles, the partial derivative of the Helmholtz free energy 
with respect to volume becomes the partial derivative of the Helmholtz 
free energy per atom, $f$,  with respect to the inverse number density. 
\begin{equation}{P=-\frac{\partial{\frac{F}{N}}}{\partial{\frac{V}{N}}}\bigg|_{T,N} = -\frac{\partial{f}}{\partial{\frac{1}{n}}}\bigg|_{T,N}}\end{equation} 
Thus, the pressure corresponds to the negative of the slope of the curve 
on a plot of the Helmholtz free energy per atom versus the inverse 
density at fixed temperature.

@@

To facilitate finding the pressure at the phase transition computationally, 
it is helpful to take different approach by considering the Gibbs free energy given by G=F+PV. 
Rearranging the terms and dividing by N gives an  expression for the 
Helmholtz free energy per atom $f$ in terms of the Gibbs free energy 
per atom  $g$ (which is the chemical potential)  
\begin{equation}\frac{F}{N} = -P\frac{V}{N}+\frac{G}{N} \end{equation}
\begin{align}f=-P\frac{1}{n}+g\end{align}
This equation has the form of a line
\begin{equation}
    y=mx+y_o
\end{equation}
and describes the black line shown in Figure~\ref{fig:MaxwellDT} with the 
pressure given by the negative of the slope of the line, and the Gibb's 
free energy per atom, or chemical potential given by the y-intercept. 
At the point of phase transition the chemical potential of the fluid is 
the same as the chemical potential of the solid, as is the pressure.
\begin{equation}
    \mu_F=\mu_S\mbox{~~~~and~~~~}P_F=P_S\mbox{~~~~at a phase transition}
\end{equation}

@@

The pressure at the phase transition can be obtained from the point of 
intersection that occurs when the chemical potential 
is plotted against the pressure for both the fluid and the 
solid at fixed temperature, as shown in Figure~\ref{fig:GibbsvsP}. 
This plot is constructed from data collected from lines tangent 
along the curves on the Helmholtz free energy per atom verses inverse 
density plots (like that of Figure~\ref{fig:MaxwellDT}) for both the 
fluid and the solid at a given temperature. 
The slope of the tangent line at a given point gives the negative of the 
pressure at that point, and the corresponding y-intercept of the tangent 
line gives the chemical potential at that point. Collecting this data 
(P,~$\mu$) at each point for the fluid and the solid respectively, 
produces the curves in Figure~\ref{fig:GibbsvsP}.

@@

Once the pressure is known at the point of phase transition, for a given 
temperature and number density, the fluid and solid number densities at 
the phase transition can be found from plots of the pressure versus the 
inverse of the number density for both the solid and the fluid, 
as shown in Figure \ref{fig:Pvsinvn}. By repeating this procedure to 
find the fluid and solid number densities at the phase transition 
for various temperatures, a Temperature-Density (T-n) phase diagram 
can be generated, similar to that shown in Figure \ref{fig:T-n_Diagram}, 
but without the gray hump characteristic of fluids that have 
attractive interatomic potentials, and which represents the region where
vapor coexists with liquid as evidenced by an interface.

@@

Equation~\ref{eq:Fexcess-simplified} is difficult to solve in general, 
but it can be solved for low densities, and this provides a way to check
a more general solution.
At low densities, it is not likely to find two or more atoms in close 
proximity, and so only interactions between individual pairs of spheres, 
such as those shown in green in Figure~\ref{fig:diagrammic}, are considered.
Each pair can then be described by a function of the distance between the 
two atoms in the pair. 
With a careful counting of the resulting number of such pairs, the excess 
free energy can then be written~\cite{schroeder} as
\begin{equation}\label{eq:Fexcess}{\frac{F_{excess}}{N}=k_BTn\left(-\frac{1}{2}\int{f(r)}{d\vec{r}}\right)}\end{equation} 
in the thermodynamic limit of large N. 
The quantity in parenthesis can be recognized 
as the second virial expansion coefficient typically called $B_2$. 
It is important that an expression for the excess Helmholtz free energy 
correctly reduces to this result in the low-density limit. 
Outside of the low density limit, it is quite difficult to evaluate the 
excess Helmholtz free energy given in Equation~\ref{eq:Fexcess-simplified}. 
SFMT, discussed in the next section, provides a way around this problem
for some types of fluids.

@@

    f(r) = -\Theta(r-(R_i + R_j))

@@

where $r=|\vec r_i - \vec r_j|$ is the distance between the centers of two spheres.

@@

$V(r)$

@@

     f(r)=\exp^{-\frac{V(r)}{k_{B}T}}-1

@@

    V(r)=\left\{ \begin{array}{rc} \infty & 0<r \leq R_i+R_j \\ 0  & r> R_i+R_j \end{array}\right.

@@

of two hard spheres added together. For a mixture of hard spheres of radius $R_a$ and $R_b$

@@

       \int_{Volume} \Theta(r-(R_i+R_j))~d\vec r &=\frac{4\pi}{3}(R_i+R_j)^3  \label{exclusion-volume}\\

@@

       &= V_i(1) + S_iR_j + R_iS_j+ V_j(1)

@@

        V &= \int_{Volume} \Theta(r-R)d\vec r \\
        S &= \int_{Volume} \delta(r-R) d\vec r \\
        R &= \int_{Volume} \frac{\delta(r-R)}{4\pi r} d\vec r \\
                1 &= \int_{Volume} \frac{\delta(r-R)}{4\pi r^2} d\vec r   

@@

Limiting this discussion to a homogeneous hard-sphere fluid composed of hard spheres that have the same
radius R, the lowest order term of the virial expansion of the excess Helmholtz free energy which is given in Equation~\ref{eq:Fexcess} 
can now be written as

@@

      F_{excess} &= k_BTn^2\left(-\frac{1}{2}\int{f(r)}{d\vec{r}}\right)V   \\
                 &= k_BTn^2\left(\int \frac{\Theta(r-2R)}{2}~d\vec r\right)V  \\
                 &= k_BTn^2(2V(1) + 2SR)V  \\
                 &= k_BT\left[\left(\int n\Theta(r-R)d\vec r\right)\left(\int n\frac{\delta(r-R)}{4\pi R^2} d\vec r \right)\color{white} \right]   \nonumber \\
                 &+ \color{white}\left[\color{black}\left(\int n\delta(r-R) d\vec r\right)\left(\int n\frac{\delta(r-R)}{4\pi R} d\vec r \right) \right]\color{black}V   \\          
                 &= k_BT(n_3n_0 + n_2n_1)V \label{Fexlowden-homo} \\
                 &= k_BT(\Phi_1 + \Phi_2)V \\
                 &= k_BT\Phi(\{n_\alpha\})V

@@

The quantity $\Phi$ has dimensions of $V^{-1}$ and is an energy density 
with a dimensionless energy composed 
of the sum of $\Phi_1$ and $\Phi_2$ which are defined as

@@

      \Phi_1 = n_3n_0  \\
      \Phi_2 = n_1n_2

@@

where $n_i$ are weighted densities of the form

@@

     n_\alpha=\int nw_\alpha(r) d\vec r

@@

with weight functions defined as
      w_3 &= \Theta(r-R)  \label{scalor-weight3} \\
      w_2 &= \delta(r-R)  \\
      w_1 &= \frac{w_2}{4\pi r}  \\

>>>>

Classically, fluids have been modeled by a collection of small, 
impenetrable spheres colliding with one another like billiard balls. 
In the simplest case, the spheres are treated as having only a repulsive 
potential. This model, called the hard-sphere fluid, has served as the 
primary reference system for fluids. This is due to the fact that the 
short-range repulsion between atoms is much stronger than their long-range 
attraction, and so the repulsive potential, which is descriptive of 
hard spheres, dominates the interactions between atoms.

In more recent times, other model fluids have been developed.
Zwanzig introduced the square-well 
fluid in 1954 which is a perturbation of the hard-sphere fluid created by adding 
a hard-edge short-range attractive potential to the repulsive potential of the 
hard-sphere fluid~\cite{ZwanzigSqrWell}. In 1964 Rowlinson studied 
spheres with softly fading short-range repulsive potentials which he modeled by making
the radius of the spheres temperature dependent~\cite{rowlinson1964statistical}. 
This reflects the idea that soft-spheres penetrate more deeply into 
each other when they collide at higher temperatures. 
In 1967 Barker and Henderson combined the perturbation approach of Zwanzig 
with Rowlinson's idea of using temperature-dependent radii to model spheres 
with soft repulsive potentials using hard-spheres~\cite{barker1967perturbation}. 
In this model, 
%a soft-sphere fluid is modeled by using a hard-sphere fluid with 
hard spheres are given larger radii 
at low temperatures, and smaller radii at high temperatures. 
In 1971 Weeks, 
Chandler, and Anderson used temperature and density dependent hard-sphere 
radii to model spheres with a more realistic softly fading repulsive 
potential similar to that between neutral atoms
%portion of the Leonard-Jones potential
which is known as the Weeks-Chandler-Anderson model fluid~\cite{andersen1971relationship} . 
Models continue to be developed to characterize fluids with particular 
properties, or to address particular problems like surface tension, 
interfaces, adsorption, wetting and confinement.
A review of some model fluids and problems that have been
studied more recently is given by L\"{o}wen~\cite{lowen}.

While models for particular fluids were being created, powerful methods 
to predict the properties of fluids were also being developed. 
In the late 1950s, Scaled Particle Theory (SPT)~\cite{ReissSPT} was 
introduced from which an equation of state, and an expression for the 
excess Helmholtz free energy of a homogeneous hard-sphere fluid can be 
obtained~\cite{Hansen}. 
In 1958, Percus-Yevick introduced a method to describe the thermodynamics 
of the hard-sphere fluid based on an approximate closure to the 
Ornstein-Zernike integral equation~\cite{PercusYevick}.
%Percus and Yevick took another approach to the hard-sphere fluid.
In 1961 Morita and Hiroike introduced a density functional 
approach~\cite{MoritaDFT} to finding the equilibrium Grand and Helmholtz free energies  
%, and the corresponding equilibrium number density profile 
of a classical inhomogeneous fluid. 
Density functional theory was further developed for inhomogeneous classical fluids
by Stillinger and Buff (1962)~\cite{stillingerandbuff}, Lebowitz and 
Percus (1963)~\cite{lobowitzandpercus}, and Mermin (1965)~\cite{merium} amoung others. A review
of the development of cDFT is found in Fundamentals of Inhomogeneous fluids~\cite{inhomogeneousfluids}.

To use classical Density Functional Theory (cDFT), an expression for the free energy 
as a functional of a spatially varying average single-particle number density 
%that varies spatially 
is constructed for a particular fluid.
%DFT Evans 1979
%Led by the consistency between results from SPT and Percus-Yevick, 
In 1989 Rosenfeld introduced Fundamental Measure Theory (FMT)~\cite{rosenfeld1989,rosenfeld1988} 
to form a functional for the hard-sphere fluid based on the idea that 
%the lowest term in the virial expansion of the excess chemical potential 
the lowest order term in the virial expansion of the excess chemical 
potential of a hard-sphere fluid can 
be expressed in terms of a weighted density using a weight function 
%using a weight 
%function 
of the form of 
a backward Heaviside step function
%\begin{equation}
   %F_{ex}=k_BT\int_{volume} n(\vec r) \theta(|\vec r - \vec r'|-(\vec R_i + \vec R_j)) d\vec r
%\end{equation}
that can itself be expressed in terms of a sum of convolutions involving 
a set of weight functions related to the geometry of a sphere. 
For two spheres with radii, 
$R_i$  and $R_j$, the convolution is written as~\cite{rosenfeld1989}
\begin{displaymath}
   \theta(|\vec r - \vec r'|-(R_i + R_j)) = w_i^{(3)} \otimes w_j^{(0)} + w_j^{(3)} \otimes w_i^{(0)} + w_i^{(1)} \otimes w_j^{(2)} + w_j^{(1)} \otimes w_i^{(2)}\end{displaymath}
\begin{equation}   + \vec w_i^{(1)} \otimes \vec w_j^{(2)} + \vec w_j^{(1)} \otimes \vec w_i^{(2)}
\end{equation}
FMT gives the same result for the excess Helmholtz free energy of the hard-sphere fluid as 
SPT does, from which it can be constructed~\cite{rosenfeld1989,Hansen,santos2012phi3},
and for the direct correlation function approximation given by Percus-Yevick. 
However, it did not predict freezing due to one term in the expression that
diverged when the fluid was highly confined.  
Tarazona later solved this problem by introducing a tensor version of 
one of the energy density terms 
in the functional~\cite{tarazonaphi3, santos2012phi3} 
which took into account the non-spherical, and thus 
non-symmetric, shape of the cavity in which the fluid is confined 
while also yielding the correct excess free energy for the fluid as its 
confinement is reduced from a three-dimensional volume (as in a bulk fluid), 
to a two-dimensional plane (as in a thin-film), to a one-dimensional 
line (as in a fluid confined to a capillary tube), to a zero-dimensional 
dot, as in an atom located at a lattice site within a crystal~\cite{rothreview}.

In 1999 Schmidt adopted FMT for spheres with soft 
interatomic potential energies
(which rise to infinity over a distance, rather than instantaneously 
like the potential energy between hard spheres)
forming Soft Fundamental Measure Theory (SFMT)~\cite{schmidt1999density}. 
%by replacing $w^{(3)}$ with a weight function that would result in  
%soft interatomic potentials (which rise to infinity over a distance, 
%rather than instantaneously).
Schmidt applied his Soft Fundamental Measure Theory to various fluids 
with soft, short-range repulsive and attractive potentials, and developed 
an error function model 
%based on a Gaussian weight function 
for spheres with steep, short-range soft repulsive 
potentials~\cite{schmidt2000fluid}.
%, similar to that exhibited by a WCA fluid. 
For some more recent developments, see the review of new developments in 
classical Density Functional Theory by Evans ~\cite{evansreview} (2016), 
Theory of Simple Fluids by Hansen and McDonald~\cite{Hansen} (2013), 
and a generalization of the Barker-Henderson model to inhomogeneous 
fluids by Tschopp and coworkers~\cite{BHextended} (2020).

Classical density functional theory has also been applied 
to phase transitions, and freezing, but not until recently has there
been good progress. Ramakrishnan and Yussouff studied freezing by viewing a 
crystal as a highly inhomogeneous liquid~\cite{ramakrishnanandyussouff} (1979). 
Oettel and fellow researchers studied hard-sphere crystals with fully developed FMT~\cite{Oettel2010} (2010). 
Mortazavifar and Oettel studied the fluid and crystal phases of the 
Asakura-Oosawa model fluid with FMT~\cite{asakuraoosawa} (2016). 
Lutsko and Lam studied studied crystallization of the Leonard-Jones 
fluid~\cite{lutsko} (2018), and Lutsko and Schoonen looked at cDFT applied 
to the solid state~\cite{lutskoschoonen2020} (2020).
In all these studies the hard-sphere fluid has been used as the reference 
system to which other terms are added to create 
the full functional for the fluid being studied.

In this paper, classical Density Functional Theory will be used to predict 
the freezing of a Weeks-Chandler-Anderson model fluid directly by applying 
SFMT with Schmidt's proposed error function model, 
in which we have incorporated our temperature dependent parameters to 
fit the functional to the soft-sphere WCA fluid. 
%form the functional for the excess Helmholtz free energy. 
%for generating the weight functions, and Tarazona's tensor functional. 
Results will be compared to those generated from Monte-Carlo simulations.

@@

The repulsive portion of the Lennard-Jones potential forms the potential for a 
Weeks-Chandler-Anderson (WCA) model fluid~\cite{andersen1971relationship},
shown in Figure \ref{fig:WCA_potential}. 
The WCA potential is described by the equation 
\begin{align} \label{eq:VWCA}
    V_{WCA}=\left\{\begin{array}{rcl} {4\epsilon{\left[\left(\frac{\sigma}{r}\right)^{12} - \left(\frac{\sigma}{r}\right)^6 \right]}+\epsilon} & \mbox{for} & 0<r<{2R} \\ 0 & \mbox{for} & r>2R \end{array}\right.
\end{align} 
The WCA potential decreases from as the distance between the 
centers of the atoms r increases, and becomes, and remains, essentially 
zero when r=$2^\frac{1}{6}\sigma=2R\approx{1.122}\sigma$ where R is the 
radius of the sphere. 
As with the Lenard-Jones potential, the repulsive WCA potential increases 
rapidly and approaches infinity as the two spheres begin to overlap. This 
is a more realistic repulsive potential than that of the hard-sphere model 
which goes instantly to infinity when the two spheres make contact (at r=2R), 
and is zero otherwise. Because the WCA spheres have no hard boundary in 
their interatomic potential, they are called ``soft spheres'' and their 
gradually fading potential makes them slightly ``squishy'', as pictured in
Figure~\ref{fig:TwoSpheres}.

@@

A phase transition occurs along a phase boundary line on a phase diagram
such as can be seen in the pressure-temperature (P-T) phase diagram
shown in Figure~\ref{fig:P-T_Diagram}. 
%
%Points along the phase boundary that separates the solid and the fluid phases 
%indicate a phase transition where the solid and fluid phases coexist, 
%and show that both the solid and the fluid have the same temperature (necessary 
%for thermal equilibrium)
%and the same pressure (necessary for mechanical 
%equilibrium)at a phase transition, and the same chemical potential 
%(necessary for diffusive equilibrium). 
%
This phase diagram is 
representative of some substance 
which exhibits both attraction and repulsion. However, since 
the atoms in a WCA fluid do not exhibit attraction there will not be a 
dividing line between the gas and liquid phases, there will just simply
be a fluid.
The phase diagram indicates the equilibrium state a substance will be in at
different temperatures and pressures.
At a phase transition, both the fluid and the solid have 
the same temperature (necessary 
for thermal equilibrium), the same pressure (necessary for mechanical 
equilibrium) and the same chemical potential (necessary for diffusive equilibrium).
%The temperature , pressure, and chemical potential will be the same for both
%the fluid and the solid at a fluid-solid phase transition.

The pressure can be related to the Helmholtz free energy F, which is minimized 
for a system at fixed temperature, volume, and number of particles as the 
entropy of the system and its surroundings added together is maximized. The negative of 
the partial derivative of the Helmholtz free energy with respect to the volume V 
at a fixed temperature T and number of particles N, equals the pressure. 
\begin{equation}{P=-\frac{\partial{F}}{\partial{V}}\bigg|_{T,N}}\end{equation}
\noindent Dividing both the Helmholtz free energy and the volume by the 
number of particles, the partial derivative of the Helmholtz free energy 
with respect to volume becomes the partial derivative of the Helmholtz 
free energy per atom, $f$,  with respect to the inverse number density. 
\begin{equation}{P=-\frac{\partial{\frac{F}{N}}}{\partial{\frac{V}{N}}}\bigg|_{T,N} = -\frac{\partial{f}}{\partial{\frac{1}{n}}}\bigg|_{T,N}}\end{equation} 
Thus, the pressure corresponds to the negative of the slope of the curve 
on a plot of the Helmholtz free energy per atom versus the inverse 
density at fixed temperature.

@@

To facilitate finding the pressure at the phase transition computationally, 
it is helpful to take different approach by considering the Gibbs free energy given by G=F+PV. 
Rearranging the terms and dividing by N gives an  expression for the 
Helmholtz free energy per atom $f$ in terms of the Gibbs free energy 
per atom  $g$ (which is the chemical potential)  
\begin{equation}\frac{F}{N} = -P\frac{V}{N}+\frac{G}{N} \end{equation}
\begin{align}f=-P\frac{1}{n}+g\end{align}
This equation has the form of a line
\begin{equation}
    y=mx+y_o
\end{equation}
and describes the black line shown in Figure~\ref{fig:MaxwellDT} where the 
pressure is given by the negative of the slope of the line, and the Gibb's 
free energy per atom, or chemical potential is given by the y-intercept. 
At the phase transition the chemical potential of the fluid is 
the same as the chemical potential of the solid, as is the pressure.
\begin{equation}
    \mu_F=\mu_S\mbox{~~~~and~~~~}P_F=P_S\mbox{~~~~at a phase transition}
\end{equation}

@@

The pressure at the phase transition can be obtained from the point of 
intersection that occurs when the chemical potential 
is plotted against the pressure for both the fluid and the 
solid at fixed temperature, as shown in Figure~\ref{fig:GibbsvsP}. 
This plot is constructed from data collected from lines tangent 
along the curves on Helmholtz free energy per atom verses inverse 
density plots (like that of Figure~\ref{fig:MaxwellDT}) for both the 
fluid and the solid at a given temperature where 
the slope of the tangent line at a given point gives the negative of the 
pressure, and the corresponding y-intercept of the tangent 
line gives the chemical potential at a given point. 
%Collecting this data 
%(P,~$\mu$) at each point for the fluid and the solid respectively, 
%produces the curves in Figure~\ref{fig:GibbsvsP}.

@@

Once the pressure is known at the point of phase transition, for a given 
temperature and number density, the fluid and solid number densities at 
the phase transition can be found from plots of the pressure versus the 
inverse of the number density for both the solid and the fluid, 
as shown in Figure \ref{fig:Pvsinvn}. By repeating this procedure to 
find the fluid and solid number densities at the phase transition 
for various temperatures, a temperature-density (T-n) phase diagram 
can be generated, similar to that shown in Figure \ref{fig:T-n_Diagram}. 
However, for a WCA fluid there will be no gray hump which is characteristic 
of fluids that have attractive interatomic potentials, and which display
an interface between vapor and liquid at temperatures below the critical point.

@@

Equation~\ref{eq:Fexcess-simplified} is difficult to solve in general, 
but it can be solved for low densities, and this provides a way to check
a more general solution.
At low densities, it is not likely to find two or more atoms in close 
proximity, and so only interactions between individual pairs of spheres, 
such as those shown in green in Figure~\ref{fig:diagrammic}, are considered.
Each pair can then be described by a function of the distance between the 
two atoms in the pair. 
With a careful counting of the resulting number of such pairs, the excess 
free energy can then be written~\cite{schroeder} as
\begin{equation}\label{eq:Fexcess}{\frac{F_{excess}}{N}\approx k_BTn\left(-\frac{1}{2}\int{f(r)}{d\vec{r}}\right)}\end{equation} 
in the thermodynamic limit of large N. 
The quantity in parenthesis can be recognized 
as the second virial expansion coefficient typically called $B_2$. 
It is important that an expression for the excess Helmholtz free energy 
correctly reduces to this result in the low-density limit. 
Outside of the low density limit, it is quite difficult to evaluate the 
excess Helmholtz free energy given in Equation~\ref{eq:Fexcess-simplified}. 
SFMT, discussed in the next section, provides a way around this problem
for some types of fluids.

@@

    f(r_{ij}) = -\Theta(r_{ij}-(R_i + R_j))

@@

where $r_{ij}=|\vec r_i - \vec r_j|$ is the distance between the centers of two spheres, 
and $\Theta$ is 1 for $r_{ij}<R_i+R_j$ and 0 for $r_{ij}>R_i+R_j$.

@@

$V(r_{ij})$

@@

     f(r_{ij})=\exp^{-\frac{V(r_{ij})}{k_{B}T}}-1

@@

    V(r_{ij})=\left\{ \begin{array}{rc} \infty & 0<r_{ij} \leq R_i+R_j \\ 0  & r_{ij}> R_i+R_j \end{array}\right.

@@

of two hard spheres added together. For a mixture of hard spheres of radius $R_i$ and $R_j$

@@

       \int_{Volume} \Theta(r_{ij}-(R_i+R_j))~d\vec r_{ij} &=\frac{4\pi}{3}(R_i+R_j)^3  \label{exclusion-volume}\\

@@

       &= V_i(1) + S_iR_j + R_iS_j+ V_j(1)  \label{VSR1sum}

@@

        V &= \int_{Volume} \Theta(r_{ij}-R)d\vec r_{ij} \\
        S &= \int_{Volume} \delta(r_{ij} -R) d\vec r_{ij}  \\
        R &= \int_{Volume} \frac{\delta(r_{ij} -R)}{4\pi r_{ij}} d\vec r_{ij} \\
        1 &= \int_{Volume} \frac{\delta(r_{ij} -R)}{4\pi r_{ij}^2} d\vec r_{ij}    

@@

Rosenfeld also recognized that the Heaviside function can be written as 

@@

    f(r_{ij})=-\Theta(r_{ij}-2R)= -2(w_3 \otimes w_0 + w_2 \otimes w_1 + \vec{w_2} \otimes \vec{w_1})

@@

where the weights are  

@@

      w_3 &= \Theta(r_{ij}-R)  \label{scalar-weight3} \\
      w_2 &= \delta(r_{ij}-R)  \\
      w_1 &= \frac{w_2}{4\pi r_{ij}}  \\
      w_0 &= \frac{w_2}{4\pi r_{ij}^2} \label{scalar-weight0}

@@

and the two vector weights are

@@

      \vec {w}_2 &= w_2~\frac{\vec r_i-\vec r_j}{r_{ij}} \\
      \vec {w}_1 &= w_1~\frac{\vec r_i-\vec r_j}{r_{ij}}

@@

which are needed for the relation to be satisfied.
%%SAVE the following - connects to convolution! ------------------------
%Rewriting the first term V(1) in Equation~\ref{VSR1sum} gives
